---
title: ' Using exponential distribution in R to validate the Central Limit Theorem'
author: "Avranil Ghosh"
date: "Sunday, January 25, 2015"
output: word_document
---

## Overview:

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be 
simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 
1/lambda. We set lambda = 0.2 for all of the simulations.

### Analysing sample mean and comparing it to the theoretical mean of the distribution.

Here we plot the histogram and density distribution of averages of 40 random exponentials simulated over 1000 times.

```{r meansim,echo=FALSE,cache=FALSE}

means = NULL
for (i in 1 : 1000) 
means = c(means, mean(rexp(40, 0.2)))
mmns<-mean(means)
mth<-1/0.2
dty <- density(means)

par(mfrow = c(1, 2))
hist(means)
plot(dty$x, dty$y, xlab = "", ylab = "density", type = "n")
lines(dty$x, dty$y, lwd = 2)
title("Distribution of means")
abline(v=mmns, col="purple")
abline(v=mth, col="blue")

```

The above figures illustrate that the mean of distribution of a large collection of averages of 40 exponentials that comes out to be `r mmns` is very close 
to the actual population theoritical mean which is `r mth`. We can conclude that the average limits to what it is estimating  - the population mean.


### Analysing sample variance and comparing it to the theoretical variance of the distribution.

Here we plot the histogram and density distribution of variances of 40 random exponentials simulated over 1000 times.

```{r sdsim,echo=FALSE,cache=FALSE}

sdev = NULL
for (i in 1 : 1000) 
sdev = c(sdev, sd(rexp(40, 0.2)))
variance<-sdev^2
mvar<-mean(variance)
sdevth<-1/0.2
varth <- sdevth^2
dtyvar <- density(variance)


par(mfrow = c(1, 2))
hist(variance)
plot(dtyvar$x, dtyvar$y, xlab = "", ylab = "density", type = "n")
lines(dtyvar$x, dtyvar$y, lwd = 2)
title("Distribution of Variance")
abline(v=mvar, col="purple")
abline(v=varth, col="blue")

```

The above figures illustrate that the mean of distribution of a large collection of variances of 40 exponentials that comes out to be `r mvar` is very close 
to the actual population theoritical variance which is `r varth`. 


### Analysing the nature of the distribution.

Here we plot the histogram of 1000 random exponential variables and compare it with the histogram of the averages of 40 random exponentials simulated over 
1000 times.

```{r dist,echo=FALSE,cache=FALSE}

exponentials<-rexp(1000, 0.2)
par(mfrow = c(1, 2))
hist(exponentials)
hist(means)

```

Here we see that although the original distribution of random exponentials is not normal, however the distribution of mean of averages assumes a normal bell 
shapped distribution. The same is true for the distribution of mean of variances.


